The five pillars of computational reproducibility: Bioinformatics and beyond

### 2023-06-06 manuscript/5pillars.Rmd
#13 Have added excel date info and refs

For a bioinformatics project to be reproducible, it needs to be enshrined by code and include all
instructions, from inspecting the raw data to generating the outputs that form the conclusions of the
study.
This automated process removes the need for any manual steps, which are not only time-consuming but
also prone to errors.
Without an end-to-end automated process, most of the other best practices described here are not
possible.
Scripted workflows, although not always free of errors, enable better auditing and easier reproduction
which would be difficult for graphical tools like spreadsheets or webtools.
Indeed, spreadsheets are prone to data entry, manipulation and formula errors @Panko1996-ko, and are
somewhat overused in the life sciences.
While webtools are valuable for data exploration, there are worries that they undermine
reproducibility for the sake of convenience [@Wijesooriya2022-ly].
Transferring data between compute platforms is also discouraged.
For example, having workflows that involve combinations of web-based and scripted tools require
data transfer steps which are inefficient and error-prone.

>>>>

For a bioinformatics project to be reproducible, it needs to be enshrined by code and include all
instructions, from inspecting the raw data to generating the outputs that form the conclusions of the
study.
This automated process removes the need for any manual steps, which are not only time-consuming but
also prone to errors.
Without an end-to-end automated process, most of the other best practices described here are not
possible.
Scripted workflows, although not always free of errors, enable better auditing and easier reproduction
which would be difficult for graphical tools like spreadsheets or webtools.
Indeed, spreadsheets are prone to data entry, manipulation and formula errors @Panko1996-ko, and are
somewhat overused and misused in the life sciences, leading to problems like inadvertant conversion
of gene names to dates in databases and supplementary Excel files [@Zeeberg2004-js;
@Ziemann2016-pv; @Abeysooriya2021-uz].
While webtools are valuable for data exploration, there are worries that they undermine
reproducibility for the sake of convenience [@Wijesooriya2022-ly].
Transferring data between compute platforms is also discouraged.
For example, having workflows that involve combinations of web-based and scripted tools require
data transfer steps which are inefficient and error-prone.
